---
title: "11_8_holly"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
options(stringsAsFactors=F)
```


Load data for Mandarin corpus
```{r echo=FALSE, message=FALSE, warning=FALSE}
mandarin <- read_csv('/Users/Yuqi/Desktop/english-l2-learner-corpus/cleaned_csv/mandarin_L1_nltk.csv')
head(mandarin)
count_uniq_words_mandarin <- read_csv('/Users/Yuqi/Desktop/english-l2-learner-corpus/cleaned_csv/mandarin_l1_lemmatizer_column.csv')
head(count_uniq_words_mandarin)

combined_mandarin <- inner_join(mandarin, count_uniq_words_mandarin, by=c("Filename"="file_id"))
head(combined_mandarin)
```

Load data for Spanish corpus
```{r echo=FALSE, message=FALSE, warning=FALSE}
spanish <- read_csv('/Users/Yuqi/Desktop/english-l2-learner-corpus/cleaned_csv/spanish_L1_nltk.csv')
head(spanish)
count_uniq_words_spanish <- read_csv('/Users/Yuqi/Desktop/english-l2-learner-corpus/cleaned_csv/spanish_l1_lemmatizer_column.csv')
head(count_uniq_words_spanish)

combined_spanish <- inner_join(spanish, count_uniq_words_spanish, by=c("Filename"="file_id"))
head(combined_spanish)
```


Higher number of unique words do somewhat lead to higher score level

Mandarin
```{r echo=FALSE, message=FALSE, warning=FALSE}
num_uniq_word_by_score_mandarin <- combined_mandarin %>%
  group_by(Score_Level) %>%
  summarize(median=median(log(num_uniq_word)))
head(num_uniq_word_by_score_mandarin)

combined_mandarin %>%
  ggplot(aes(x=log(num_uniq_word), color=Score_Level, fill=Score_Level)) +
  geom_density(alpha=0.3,size=1)+ 
  xlim(4,6)+
  geom_vline(data=num_uniq_word_by_score_mandarin, aes(xintercept = median, color = Score_Level), size=1.5)+ 
  labs(x= "Logged number of unique words in an essay")
  theme(legend.position="bottom")
``` 

Spanish 
```{r echo=FALSE, message=FALSE, warning=FALSE}
num_uniq_word_by_score_spanish <- combined_spanish %>%
  group_by(Score_Level) %>%
  summarize(median=median(log(num_uniq_word)))
head(num_uniq_word_by_score_spanish)

combined_spanish %>%
  ggplot(aes(x=log(num_uniq_word), color=Score_Level, fill=Score_Level)) +
  geom_density(alpha=0.3,size=1)+ 
  geom_vline(data=num_uniq_word_by_score_spanish, aes(xintercept = median, color = Score_Level), size=1.5)+ 
  labs(x= "Logged number of unique words in an essay")
``` 


```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(lm(combined_mandarin$noun_per_sentence~combined_mandarin$num_uniq_word))
ggplot(combined_mandarin, aes(x=num_uniq_word, y=noun_per_sentence/verb_per_sentence)) + geom_point() + stat_smooth(method='lm')
```


Number of nouns and number of "the" has quite a positive correlation , which is not surprising because most of nouns need some sort of determiner in English. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(combined_mandarin, aes(x=noun_count, y=the_count, color=Score_Level)) + geom_point() + stat_smooth(method='lm')
``` 


Plotting num uniq word vs. the freq
```{r echo=FALSE, message=FALSE, warning=FALSE}
combined_mandarin %>%
  ggplot(aes(x=num_uniq_word, y=log(the_freq), color=Score_Level)) +
  geom_point() +
  stat_smooth(method='lm') +
  labs(x= "Number of unique words in an essay",
        subtitle="Number of uniq words vs. Log frequency of \"the\", Mandarin corpus")

combined_spanish %>%
  ggplot(aes(x=num_uniq_word, y=log(the_freq), color=Score_Level)) +
  geom_point() +
  stat_smooth(method='lm') +
  labs(x= "Number of unique words in an essay",
        subtitle="Number of uniq words vs. Log frequency of \"the\", Spanish corpus")
```


"The" frequency plotted against number of noun per sentence 
--- different score levels seem to have similar slope (all are pretty weak though)
```{r echo=FALSE, message=FALSE, warning=FALSE}
combined_mandarin %>%
  ggplot(aes(x=noun_per_sentence, y=log(the_freq), color=Score_Level)) +
  geom_point() +
  xlim(0, 15) + 
  stat_smooth(method='lm') +
  labs(subtitle="Number of noun per sentence vs. Log frequency of \"the\", Mandarin corpus")


combined_spanish %>%
  ggplot(aes(x=noun_per_sentence, y=log(the_freq), color=Score_Level)) +
  geom_point() +
  xlim(0, 15) + 
  stat_smooth(method='lm') +
  labs(subtitle="Number of noun per sentence vs. Log frequency of \"the\", Spanish corpus")

```

Plotting a freq vs. the freq
```{r echo=FALSE, message=FALSE, warning=FALSE}
combined_mandarin %>%
  ggplot(aes(x=log(aan_freq), y=log(the_freq), color=Score_Level)) +
  geom_point() +
  stat_smooth(method='lm') +
  labs(x= "Frequency of a/an",
        subtitle="Frequency of the vs. Frequency of a/an, Mandarin corpus")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
combined_mandarin %>%
  ggplot(aes(x=num_uniq_word, y=log(the_freq)/log(aan_freq), color=Score_Level)) +
  geom_point() +
  stat_smooth(method='lm') +
  labs(subtitle="Ratio between the freq and aan freq, Mandarin corpus")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
combined_mandarin %>%
  ggplot(aes(x=num_uniq_word, y=log(aan_freq))) +
  geom_point() +
  stat_smooth() +
  labs(subtitle="Logged frequency of a/an vs. number of unique words, Mandarin corpus")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
combined_mandarin %>%
  ggplot(aes(x=num_uniq_word, y=log(the_freq))) +
  geom_point() +
  stat_smooth(method='lm') +
  labs(subtitle="Logged frequency of the vs. number of unique words, Mandarin corpus")
```



Density plot for ratio between the and a/an 
```{r}
the_aan_ratio_mandarin <- combined_mandarin %>%
  group_by(Score_Level) %>%
  summarize(median=median(log(the_count/aan_count), na.rm=TRUE))
head(the_aan_ratio_mandarin)

combined_mandarin %>%
  ggplot(aes(x=log(the_count/aan_count), color=Score_Level, fill=Score_Level)) +
  geom_density(alpha=0.3,size=1)+ 
  geom_vline(data=the_aan_ratio_mandarin, aes(xintercept = median, color = Score_Level), size=1)+ 
  xlim(0, 5) + 
  labs(x= "Log-transformed ratio between number of \"the\" and number of \"a/an\", Mandarin corpus")+
  theme(legend.position="bottom")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(combined_mandarin, aes(x=num_uniq_word)) +
  geom_point(aes(y=log(the_freq)), col="blue") +
  geom_point(aes(y=log(aan_freq)), col="red") + 
  geom_smooth(aes(y=log(the_freq)), col="green") +
  geom_smooth(aes(y=log(aan_freq)), col="green") +
  scale_y_continuous(name="the", sec.axis=sec_axis(~ ., name="a/an")) +
  theme(
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red")
  ) + 
  labs(subtitle="Log frequency of \"the\" and \"a/an\", Mandarin corpus")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
combined_mandarin %>%
  ggplot(aes(x=num_uniq_word, y=log(the_count/(this_count+these_count+that_count+those_count)))) +
  geom_point() +
  stat_smooth() +
  labs(x= "Number of unique words", y="Logged ratio between 'the' and demonstratives")
```
```{r}
the_demo_ratio_mandarin <- combined_mandarin %>%
  group_by(Score_Level) %>%
  summarize(average=mean(the_count/(this_count+these_count+that_count+those_count+1), na.rm=TRUE))%>%
  mutate(average=log(average))
head(the_demo_ratio_mandarin)

combined_mandarin %>%
  ggplot(aes(x=log(the_count/(this_count+these_count+that_count+those_count)), color=Score_Level, fill=Score_Level)) +
  geom_density(alpha=0.3,size=1)+ 
  geom_vline(data=the_demo_ratio_mandarin, aes(xintercept = average, color = Score_Level), size=1)+ 
  xlim(0, 5) + 
  labs(x= "Logged ratio between number of \"the\" and number of demonstratives, Mandarin L1 corpus")
```

```{r}
the_demo_ratio_spanish <- combined_spanish %>%
  group_by(Score_Level) %>%
  summarize(average=mean(the_count/(this_count+these_count+that_count+those_count+1), na.rm=TRUE))%>%
  mutate(average=log(average))
head(the_demo_ratio_spanish)

combined_spanish %>%
  ggplot(aes(x=log(the_count/(this_count+these_count+that_count+those_count)), color=Score_Level, fill=Score_Level)) +
  geom_density(alpha=0.3,size=1)+ 
  geom_vline(data=the_demo_ratio_spanish, aes(xintercept = average, color = Score_Level), size=1)+ 
  xlim(0, 5) + 
  labs(x= "Logged ratio between number of \"the\" and number of demonstratives, Spanish L1 corpus")
```

Plotting a freq vs. the freq
```{r echo=FALSE, message=FALSE, warning=FALSE}
combined_mandarin %>%
  ggplot(aes(x=this_freq, y=the_freq, color=Score_Level)) +
  geom_point() +
  stat_smooth(method='lm') +
  labs(x= "Frequency of this",
        subtitle="Frequency of the vs. Frequency of this, Mandarin corpus")
```

Density plot for ratio between the and demonstratives --- high and medium are almost overlapping, low is different
```{r}
combined_mandarin %>%
  ggplot(aes(x=log(the_freq)/log(this_freq), color=Score_Level, fill=Score_Level)) +
  geom_density(alpha=0.3,size=1)+ 
  #geom_vline(data=num_uniq_word_by_score_mandarin, aes(xintercept = median, color = Score_Level), size=1.5)+ 
  labs(x= "Ratio between the freq and this freq, Mandarin corpus")+
  theme(legend.position="bottom")
```

Ratio between the and other det counts
```{r}

the_other_ratio_mandarin <- combined_mandarin %>%
  group_by(Score_Level) %>%
  summarize(median=median(log(the_count/other_det_count), na.rm=TRUE))
head(the_other_ratio_mandarin)

combined_mandarin %>%
  ggplot(aes(x=log(the_count/other_det_count), color=Score_Level, fill=Score_Level)) +
  geom_density(alpha=0.3,size=1)+ 
  xlim(0, 5)+ 
  geom_vline(data=the_other_ratio_mandarin, aes(xintercept = median, color = Score_Level), size=1)+ 
  labs(x= "Logged ratio between number of \"the\" and other determiners, Mandarin corpus")+
  theme(legend.position="bottom")
```

```{r}

the_other_ratio_spanish <- combined_spanish %>%
  group_by(Score_Level) %>%
  summarize(median=mean(log(the_count/other_det_count), na.rm=TRUE))
head(the_other_ratio_spanish)

combined_spanish %>%
  ggplot(aes(x=the_count/other_det_count, color=Score_Level, fill=Score_Level)) +
  geom_density(alpha=0.3,size=1)+ 
  xlim(0, 15)+ 
  geom_vline(data=log(the_other_ratio_spanish), aes(xintercept = median, color = Score_Level), size=1)+ 
  labs(x= "Ratio between number of \"the\" and other determiners, Spanish corpus")+
  theme(legend.position="bottom")
```

Plotting this freq
```{r echo=FALSE, message=FALSE, warning=FALSE}
combined_mandarin %>%
  ggplot(aes(x=num_uniq_word, y=this_freq+that_freq+these_freq+those_freq)) +
  geom_point() +
  stat_smooth(method='lm') +
  labs(x= "Frequency of this",
        subtitle="Frequency of the vs. Frequency of this, Mandarin corpus")
```

plot 2 lines: freq of demonstratives and freq of definite articles for mandarin speakers
x axis: num unique words

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(combined_mandarin, aes(x=num_uniq_word)) +
  geom_point(aes(y=log(the_freq)), col="blue") +
  geom_point(aes(y=log(this_freq+that_freq+these_freq+those_freq)), col="red") + 
  geom_smooth(aes(y=log(the_freq)), col="green") +
  geom_smooth(aes(y=log(this_freq+that_freq+these_freq+those_freq)), col="green") +
  scale_y_continuous(name="the", sec.axis=sec_axis(~ ., name="demonstratives")) +
  theme(
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red")
  ) + 
  labs(subtitle="Log frequency of \"the\" and the 4 demonstratives, Mandarin corpus")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(combined_spanish, aes(x=num_uniq_word)) +
  geom_point(aes(y=log(the_freq)), col="blue") +
  geom_point(aes(y=log(this_freq+that_freq+these_freq+those_freq)), col="red") + 
  geom_smooth(aes(y=log(the_freq)), col="green") +
  geom_smooth(aes(y=log(this_freq+that_freq+these_freq+those_freq)), col="green") +
  scale_y_continuous(name="the", sec.axis=sec_axis(~ ., name="demonstratives")) +
  theme(
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red")
  ) + 
  labs(subtitle="Log requency of \"the\" and the 4 demonstratives, Spanish corpus")

```

Only looking at essays with high scores
```{r echo=FALSE, message=FALSE, warning=FALSE}
mandarin_high <-
  combined_mandarin %>%
  filter(Score_Level=="high") %>%
  select(Filename, Language, Score_Level, num_uniq_word, the_freq)
nrow(mandarin_high)
summary(lm(mandarin_high$the_freq~mandarin_high$num_uniq_word))

spanish_high <-
  combined_spanish %>%
  filter(Score_Level=="high") %>%
  select(Filename, Language, Score_Level, num_uniq_word, the_freq)
nrow(spanish_high)

combined_high_score <- rbind(mandarin_high, spanish_high)
nrow(combined_high_score)

combined_high_score %>%
  ggplot(aes(x=num_uniq_word, y=the_freq, color=Language)) + 
  geom_point() +
  stat_smooth(method='lm') +
  labs(x= "Number of unique words in an essay",
       y= "Proportion of \"the\" over all words in an essay", 
        subtitle="Number of uniq words vs. Frequency of \"the\", \"High\" scores from both corpora")

```

Only looking at essays with low scores
```{r echo=FALSE, message=FALSE, warning=FALSE}
mandarin_low <-
  combined_mandarin %>%
  filter(Score_Level=="low") %>%
  select(Filename, Language, Score_Level, num_uniq_word, the_freq)
nrow(mandarin_low)
summary(lm(mandarin_high$the_freq~mandarin_high$num_uniq_word))

spanish_low <-
  combined_spanish %>%
  filter(Score_Level=="low") %>%
  select(Filename, Language, Score_Level, num_uniq_word, the_freq)
nrow(spanish_low)

combined_low_score <- rbind(mandarin_low, spanish_low)
nrow(combined_low_score)

combined_low_score %>%
  ggplot(aes(x=num_uniq_word, y=the_freq, color=Language)) + 
  geom_point() +
  stat_smooth(method='lm') +
  labs(x= "Number of unique words in an essay",
       y= "Proportion of \"the\" over all words in an essay", 
        subtitle="Number of uniq words vs. Frequency of \"the\", \"Low\" scores from both corpora")

```

PCA based on sentence complexity metrics, Mandarin corpus

Could predict the score level from these factors
Probably shouldn't do PCA on the predicted values

```{r}
sent_complex_mandarin <- 
  combined_mandarin %>%
  mutate(avg_sent_len=essay_len/num_sentence) %>%
  mutate(the_aan_ratio=the_count/(aan_count + 1)) %>%
  select(noun_per_sentence, num_uniq_word, avg_sent_len, the_aan_ratio) 
  
head(sent_complex_mandarin)

(sent_complex_pca_mandarin <- 
  prcomp(sent_complex_mandarin[c('noun_per_sentence', 'num_uniq_word', 'avg_sent_len')],center=T, scale=T))

sent_complex_pca_mandarin$rotation

summary(lm(sent_complex_mandarin$the_aan_ratio
           ~ sent_complex_pca_mandarin$x[,"PC1"] +
               sent_complex_pca_mandarin$x[,"PC2"] +
               sent_complex_pca_mandarin$x[,"PC3"]))
``` 

PCA based on sentence complexity metrics, Spanish corpus
```{r}
sent_complex_spanish <- 
  combined_spanish %>%
  mutate(avg_sent_len=essay_len/num_sentence) %>%
  mutate(the_aan_ratio=the_count/(aan_count + 1)) %>%
  select(noun_per_sentence, num_uniq_word, avg_sent_len, the_aan_ratio) 
  
head(sent_complex_spanish)

(sent_complex_pca_spanish <- 
  prcomp(sent_complex_spanish[c('noun_per_sentence', 'num_uniq_word', 'avg_sent_len')],center=T, scale=T))

sent_complex_pca_spanish$rotation

summary(lm(sent_complex_spanish$the_aan_ratio
           ~ sent_complex_pca_spanish$x[,"PC1"] +
               sent_complex_pca_spanish$x[,"PC2"] +
               sent_complex_pca_spanish$x[,"PC3"]))
``` 

```{r}
sent_complex_mandarin <- 
  combined_mandarin %>%
  mutate(avg_sent_len=essay_len/num_sentence) %>%
  mutate(the_aan_ratio=the_count/(aan_count + 1)) %>%
  mutate(score_level_num=ifelse(Score_Level=="low", 1, ifelse(Score_Level=="medium", 2, 3))) %>%
  select(noun_per_sentence, num_uniq_word, avg_sent_len, Score_Level, score_level_num) 
  
head(sent_complex_mandarin)

(sent_complex_pca_mandarin <- 
  prcomp(sent_complex_mandarin[c('noun_per_sentence', 'num_uniq_word', 'avg_sent_len')],center=T, scale=T))

sent_complex_pca_mandarin$rotation

summary(lm(sent_complex_mandarin$score_level_num
           ~ sent_complex_pca_mandarin$x[,"PC1"] +
               sent_complex_pca_mandarin$x[,"PC2"] +
               sent_complex_pca_mandarin$x[,"PC3"]))
```